### 2025-02-20 manuscript/manuscript.qmd
feat(manuscript): update results following merge cc98abi

TABLE 1: Summary of data (part of `analysis/preprocessing.qmd` updated 21/1/2025)

|            | Regression                      | Classification            | Total |
|:-----------|:--------------------------------|:--------------------------|:------|
| Study N    | 22                              | 12                        | 34    |
| Model N    | 204                             | 86                        | 290   |
| Techniques | Neural Nets: 66                 | 25                        | 91    |
| Techniques | Support Vector Machines: 50     | 26                        | 76    |
| Techniques | Linear Methods: 62              | 13                        | 75    |
| Techniques | Tree-based Methods: 14          | 10                        | 34    |
| Techniques | KS, Add. & KNN: 12              | 12                        | 24    |
| Feature N  | Min=3, Md=605, Max=6670         | Min=8, Md=126, Max=8904   | NA    |
| Stimulus N | Min=20, Md=324, Max=2486        | Min=124, Md=300, Max=5192 | NA    |

Although the total number of studies meeting the criteria described in the previous section is modest (34 in total), they encompass a large array of models (290 in total) with a relatively even distribution among the three most popular techniques: Neural Nets, SVMs, and Linear Methods. The number of features and stimuli within these studies varies significantly, ranging from as few as three features [@battcock2021in] to a maximum of almost 9000 features [@zhang2023mo]. The median number of features differs between regression (605) and classification (126) studies, primarily reflecting the nature of the datasets used in each approach. The number of stimuli is typically around 300-400 (with a median of 324 for regression and 300 for classification), though there is substantial variation, with the extremes from 20 stimuli in @beveridge2018po to 5192 stimuli in @alvarez2023ri. There are also additional dimensions to consider, such as the type of cross-validation used, the music genres analyzed (whether a single genre, multiple genres, or a mix), the type of journal in which the studies were published, and the extraction tool used to extract features. However, these variables do not lend themselves to a simple summary, so we will revisit them during the interpretation and discussion stages.

@@

| Concept      | Models, obs | $r$ \[95%-CI\]          | $t$   | $p$   | $\tau^2$ | $I^2$ |
|:-------------|:------------|:------------------------|:------|:------|:---------|:------|
| Valence All  | 102,60017   | 0.583 \[0.541-0.623\]   | 21.40 | .0001 | 0.094    | 97.7% |
| Valence Max  | 22,14172    | 0.669 \[0.560-0.755\]   | 9.58  | .0001 | 0.148    | 98.4% |
| *N Features* |             |                         |       |       |          |       |
| \<18 F       | 5,3036      | 0.811 \[0.542-0.823\]   | \-    | \-    | 0.182    | 98.9% |
| 18-260 F     | 11,7042     | 0.584 \[0.534-0.931\]   | \-    | \-    | 0.133    | 97.6% |
| 260+ F       | 5,3494      | 0.710 \[0.542-0.823\]   | \-    | \-    | 0.044    | 97.2% |
| *Techniques* |             |                         |       |       |          |       |
| KS           | 1,1838      | 0.466 \[-0.634-0.942\]  | \-    | \-    | 0.0185   | 95.1% |
| LM           | 8,1762      | 0.797 \[0.614-0.899\]   | \-    | \-    | 0.1370   | 96.4% |
| SVM          | 6,4993      | 0.656 \[0.484-0.779\]   | \-    | \-     | 0.0574  | 96.9% |
| NN           | 4,2249      | 0.393 \[-0.355-0.835\]  | \-    | \-     | 0.0761  | 97.1% |
| TM           | 3,3330      | 0.750 \[0.292-0.928\]   | \-    | \-     | 0.093   | 98.5% |

<!-- table updated TE 21/1/2025, revised terms 28/1/2025 -->

@@

<!-- numbers in the text updated up to here, TE 21/1/2025 -->

@@

To gain insights into the factors contributing to the wide range of model success, we explored several ways of splitting the data. Table 2 presents two key splits: one based on the number of features used, which we hypothesized might influence model performance, and another based on the modeling techniques employed. In terms of feature sets, we categorized them into three groups: few features (<18), a large number of features (18–260), and massive feature sets (260+). Interestingly, models (4 in total) using a relatively small number of features (<18) performed best (_r_ = 0.817, 95% CI: 0.361–0.958) compared to those utilizing larger feature sets. However, it is worth noting that the models using massive feature sets (260+, 7 studies in total) also performed well (_r_ = 0.702), achieving better and more consistent results than the overall prediction rate (_r_ = 0.669). This observation is supported by the lowest heterogeneity index for the massive feature set group ($\tau^2$ = 0.029), indicating more consistent results across studies. Studies with large number of features (18-260 features, 10 studies in total) delivered the worst results, _r_ = 0.586 (95% CI: 0.360–0.748). Despite the fluctuation in the overall model accuracy between the number of features, the differences are not substantially large to pass the test of statistical significance (_Q_(2) = 3.47, _p_=.176).

<!-- numbers in the text updated up to here, TE 21/1/2025 -->

When analyzing the studies across the five modeling techniques used, the prediction do differ significantly (_Q_(4) = 45.7, _p_ < .0001). Notably, linear models (LM) and Support Vector Machines (SVM) were the most common, with 8 and 6 studies, respectively, allowing for more confident interpretations. Linear models achieved the highest prediction rate (_r_ = 0.784, 95% CI: 0.625–0.881), though this may be influenced by the smaller datasets typically used in these studies. These studies also exhibited substantially higher heterogeneity ($\tau^2$ = 0.137) compared to other techniques, where $\tau^2$ values were less than half of this. While there were only 3 studies involving tree-based model (TM), these performed well, achieving _r_ = 0.750, 95% CI: 0.292–0.928), and the relatively poor performance of the neural network (NN) models represented in four studies (_r_ = 0.340, 95% CI: -0.097–0.668) is difficult to explain without a deeper examination of the specific model architectures and the stimuli used in these studies. Kernel smoothing and KNN models (with only 1 study represented) achieved relatively poor success as well, _r_ = 0.383, 95% CI: 0.348–0.426).

We also ran the sub-grouping analyses across a combination of stimulus genres (single vs mixed) and number of the features to explore where the differences in the model prediction rates might lie. For this purpose, we grouped the studies into small single genre/multigenre studies, medium single genre/multigenre studies, and medium-large single genre/multigenre studies, and huge multigenre studies. The small single genre/multigenre studies generally performed best (_r_ = 0.836), followed by huge multi-genre studies (_r_ = 0.693) while the medium and medium-large sized studies performance was between close the overall average (medium-large, _r_ = 0.768 and medium _r_ = 0.607). The heterogeneity was lowest in the huge multigenre studies ($\tau^2$ = 0.013) and highest in the small single genre/multigenre studies ($\tau^2$ = 0.318).

@@

Moving on the arousal, we carry out the same meta-analytical analysis applying the random-effects model to arousal. Table 3 describes the broad pattern of results in tabular format, and Figure 3 illustrates the spread and heterogeneity of all studies for arousal. The overall correlation across the studies using the best performing model out of each study (Max) is 0.809 (CI95% 0.740-0.860). If we examine all the models reported in each study, the correlation drops marginally, to 0.791 (CI95% 0.770-0.810), despite this analysis includes about four times as many models as taking the best model out of each study. For arousal, even the baseline models seem to be performing on a relative high level. However, the indicators of heterogeneity are again high ($\tau^2$ = 0.141 and $I^2$=97.9%), which suggests that summary may be misleading. However, the analysis of asymmetry does not reveal significant issues (Eggers test, $\beta$ = 0.789 95%CI -4.87-6.45, _t_ = 0.273, _p_ = 0.788). If we remove the studies that are outside the 95%CI in heterogeneity, leaves this 13 studies in the summary that has _r_ = 0.826 (95%CI 0.806-0.845) with $\tau^2$ = 0.0042 and $I^2$ = 76.8%. In other words, no material difference to the results obtained with all 22 studies.

@@

| Concept      | Models, obs | $r$ \[95%-CI\]            | $t$  | $p$    | $\tau^2$ | $I^2$ |
|:-------------|:------------|:--------------------------|:-----|:-------|:---------|:------|
| Arousal All  | 102, 60017  | 0.791 \[0.770-0.810\]     | 39.9 | 0.0001 | 0.069    | 96.2% |
| Arousal Max  | 22, 14172   | 0.809 \[0.740-0.860\]     | 13.6 | 0.0001 | 0.141    | 97.9% |
| *N Features* |             |                           |      |        |          |       |
| \<18 F       | 4, 3016     | 0.817 \[0.361-0.958\]     |      |        | 0.2295   | 99.2% |
| 18-260 F     | 10, 4222    | 0.586 \[0.356-0.748\]     |      |        | 0.1542   | 96.9% |
| 260+ F       | 7, 6190     | 0.702 \[0.612-0.774\]     |      |        | 0.0285   | 96.6% |
| *Techniques* |             |                           |      |        |          |       |
| KS           | 1, 1838     | 0.819 \[0.803-0.833\]     |      |        |  -       | -     |
| LM           | 8, 1762     | 0.881 \[0.808-0.928\]     |      |        | 0.0846   | 93.3% |
| SVM          | 6, 4993     | 0.808 \[0.644-0.901\]     |      |        | 0.1125   | 98.0% |
| NN           | 4, 2249     | 0.533 \[0.328-0.691\]     |      |        | 0.0190   | 85.8% |
| TM           | 3, 3330     | 0.809 \[0.733-0.864\]     |      |        | 0.0025   | 65.4% |

<!-- numbers in the text updated up to here, TE 22/1/2025 -->

@@

Figure 3 presents (A) a forest and (B) funnel plot of the random-effects model of the best-performing models from all studies. Similarly to valence, the range of correlations is also wide for arousal, ranging from 0.35 to 0.95, with all studies demonstrating evidence of positive correlations. A mean estimate of 0.81 or higher is achieved by the majority (15 out of the 22 models). Due to large sample in most studies, the confidence intervals are narrow, although the exceptions ($N < 55$) are clearly visible [@beveridge2018po; @griffiths2021am;@koh2023me; @saizclar2022pr; @wang2021ac]. The funnel plot in panel B illustrates the limited range of the correlations and show heavy clustering at the top of the plot (studies with low standard error). There is no clear asymmetry in the plot, verified by non-significant Egger's test reported above.

The analysis of the subdivision of studies shows that there is no significant differences between the studies using different number of features (_Q_(2) = 3.47, _p_ = .176) despite the differing means (_r_ = 0.817 for studies with less than 18 features, _r_ = 0.586 for 18 to 260 features, and _r_ = 0.702 for studies utilising over 260 features). The differences in the techniques, however, does show statistically significant variance between subgroups (_Q_(4) = 50.47, _p_ < .0001). The Neural Nets (NN) achieve poor prediction of arousal (_r_ = 0.533) in comparison to other techniques. The caveat of this subgroup analysis is the small number of observations for five techniques. When other subgroupings were explored, such as the type of journal (Engineering with 13 studies vs Psychology with 9 studies), no significant differences were observed (_Q_(1) = 0.94, _p_ = .333).

{{< include _figure3.qmd >}}

>>>>

TABLE 1: Summary of data (part of `analysis/preprocessing.qmd` updated 20/2/2025)

|            | Regression                      | Classification            | Total |
|:-----------|:--------------------------------|:--------------------------|:------|
| Study N    | 22                              | 12                        | 34    |
| Model N    | 204                             | 86                        | 290   |
| Techniques | Neural Nets: 64                 | 21                        | 85    |
| Techniques | Support Vector Machines: 62     | 26                        | 88    |
| Techniques | Linear Methods: 62              | 19                        | 81    |
| Techniques | Tree-based Methods: 14          | 16                        | 30    |
| Techniques | KS, Add. & KNN:  2              | 4                         | 6     |
| Feature N  | Min=3, Md=653, Max=14460        | Min=6, Md=98, Max=8904    | NA    |
| Stimulus N | Min=20, Md=324, Max=2486        | Min=124, Md=300, Max=5192 | NA    |

Although the total number of studies meeting the criteria described in the previous section is modest (34 in total), they encompass a large array of models (290 in total) with a relatively even distribution among the three most popular techniques: Neural Nets, SVMs, and Linear Methods. The number of features and stimuli within these studies varies significantly, ranging from as few as three features [@battcock2021in] to a maximum of almost 14,500 features [@zhang2023mo]. The median number of features differs between regression (653) and classification (98) studies, primarily reflecting the nature of the datasets used in each approach. The number of stimuli is typically around 300-400 (with a median of 324 for regression and 300 for classification), though there is substantial variation, with the extremes from 20 stimuli in @beveridge2018po to 5192 stimuli in @alvarez2023ri. There are also additional dimensions to consider, such as the type of cross-validation used, the music genres analyzed (whether a single genre, multiple genres, or a mix), the type of journal in which the studies were published, and the extraction tool used to extract features. However, these variables do not lend themselves to a simple summary, so we will revisit them during the interpretation and discussion stages.

@@

| Concept      | Models, obs | $r$ \[95%-CI\]          | $t$   | $p$   | $\tau^2$ | $I^2$ |
|:-------------|:------------|:------------------------|:------|:------|:---------|:------|
| Valence All  | 102,60017   | 0.583 \[0.541-0.623\]   | 21.41 | .0001 | 0.094    | 97.7% |
| Valence Max  | 22,14172    | 0.669 \[0.560-0.755\]   | 9.58  | .0001 | 0.148    | 98.4% |
| *N Features* |             |                         |       |       |          |       |
| \<30 F       | 6,3140      | 0.766 \[0.488-0.903\]   | \-    | \-    | 0.198    | 98.6% |
| 30-300 F     | 8,4098      | 0.580 \[0.276-0.778\]   | \-    | \-    | 0.188    | 97.4% |
| 300+ F       | 8,6934      | 0.666 \[0.531-0.767\]   | \-    | \-    | 0.062    | 98.0% |
| *Techniques* |             |                         |       |       |          |       |
| LM           | 9,2457      | 0.784 \[0.652-0.870\]   | \-    | \-    | 0.1194   | 96.3% |
| SVM          | 4,5068      | 0.539 \[0.171-0.774\]   | \-    | \-    | 0.0702   | 97.1% |
| NN           | 6,3317      | 0.473 \[0.167-0.696\]   | \-    | \-    | 0.1029   | 98.2% |
| TM           | 3,3330      | 0.750 \[0.292-0.928\]   | \-    | \-    | 0.0740   | 98.8% |

<!-- table updated CA 20/2/2025, revised terms 28/1/2025 -->

@@

<!-- numbers in the text updated up to here, CA 20/2/2025 -->

@@

To gain insights into the factors contributing to the wide range of model success, we explored several ways of splitting the data. Table 2 presents two key splits: one based on the number of features used, which we hypothesized might influence model performance, and another based on the modeling techniques employed. In terms of feature sets, we categorized them into three groups: few features (<30), a large number of features (30–300), and massive feature sets (300+). Interestingly, models (6 in total) using a relatively small number of features (<30) performed best (_r_ = 0.766, 95% CI: 0.488–0.903) compared to those utilizing larger feature sets. However, it is worth noting that the models using massive feature sets (300+, 8 studies in total) also performed reasonably well (_r_ = 0.666), achieving more consistent results than the overall prediction rate (_r_ = 0.669). This observation is supported by the lowest heterogeneity index for the massive feature set group ($\tau^2$ = 0.062), indicating more consistent results across studies. Studies with large number of features (30-300 features, 8 studies in total) delivered the worst results, _r_ = 0.580 (95% CI: 0.276–0.778). Despite the fluctuation in the overall model accuracy between the number of features, the differences are not substantially large to pass the test of statistical significance (_Q_(2) = 2.03, _p_=.363).

<!-- numbers in the text updated up to here, CA 20/2/2025 -->

When analyzing the studies across the four modeling techniques used, the prediction do differ significantly (_Q_(3) = 12.43, _p_ = .0061). Notably, linear models (LM) and neural networks (NN) were the most common, with 9 and 6 studies, respectively, allowing for more confident interpretations. Linear models achieved the highest prediction rate (_r_ = 0.784, 95% CI: 0.652–0.870), though this may be influenced by the smaller datasets typically used in these studies. These studies also exhibited higher heterogeneity ($\tau^2$ = 0.119) compared to other techniques. While there were only 3 studies involving tree-based model (TM), these performed well, achieving _r_ = 0.750, 95% CI: 0.292–0.928), and the relatively poor performance of the neural network (NN) models represented in six studies (_r_ = 0.473, 95% CI: 0.167–0.696) is difficult to explain without a deeper examination of the specific model architectures and the stimuli used in these studies.

We also ran the sub-grouping analyses across stimulus genres (single vs mixed), finding no significant difference (_Q_(1) = 0.01, _p_ = .9158). Both single-genre (_r_ = 0.675 95% CI: 0.465, 0.813, _n_ = 8) and multi-genre (_r_ = 0.665, 95% CI: 0.508-0.779, _n_ = 14) achieved similar results, although multi-genre studies exhibited slightly higher heterogeneity ($\tau^2$ = 0.167) than single-genre studies ($\tau^2$ = 0.137).

@@

Moving on the arousal, we carry out the same meta-analytical analysis applying the random-effects model to arousal. Table 3 describes the broad pattern of results in tabular format, and Figure 3 illustrates the spread and heterogeneity of all studies for arousal. The overall correlation across the studies using the best performing model out of each study (Max) is 0.809 (CI95% 0.740-0.860). If we examine all the models reported in each study, the correlation drops marginally, to 0.791 (CI95% 0.770-0.810), despite this analysis including about four times as many models as taking the best model out of each study. For arousal, even the baseline models seem to perform at a relatively high level. However, the indicators of heterogeneity are again high ($\tau^2$ = 0.141 and $I^2$=97.9%), which suggests that summary may be misleading. However, the analysis of asymmetry does not reveal significant issues (Eggers test, $\beta$ = 0.789 95%CI -4.87-6.45, _t_ = 0.273, _p_ = 0.788). If we remove the studies that are outside the 95%CI in heterogeneity, leaves this 13 studies in the summary that has _r_ = 0.826 (95%CI 0.806-0.845) with $\tau^2$ = 0.0042 and $I^2$ = 76.8%. In other words, no material difference to the results obtained with all 22 studies.

@@

| Concept      | Models, obs | $r$ \[95%-CI\]            | $t$  | $p$    | $\tau^2$ | $I^2$ |
|:-------------|:------------|:--------------------------|:-----|:-------|:---------|:------|
| Arousal All  | 102, 60017  | 0.791 \[0.770-0.810\]     | 39.9 | 0.0001 | 0.069    | 96.2% |
| Arousal Max  | 22, 14172   | 0.809 \[0.740-0.860\]     | 13.6 | 0.0001 | 0.141    | 97.9% |
| *N Features* |             |                           |      |        |          |       |
| \<30 F       | 6, 3140     | 0.885 \[0.782-0.940\]     |      |        | 0.0948   | 93.5% |
| 30-300 F     | 8, 4098     | 0.735 \[0.501-0.868\]     |      |        | 0.1971   | 98.2% |
| 300+ F       | 8, 6934     | 0.804 \[0.716-0.867\]     |      |        | 0.0612   | 97.4% |
| *Techniques* |             |                           |      |        |          |       |
| LM           | 8, 1713     | 0.882 \[0.809-0.928\]     |      |        | 0.0846   | 93.3% |
| SVM          | 5, 5812     | 0.796 \[0.559-0.913\]     |      |        | 0.1325   | 98.3% |
| NN           | 6, 3317     | 0.660 \[0.395-0.823\]     |      |        | 0.1209   | 98.1% |
| TM           | 3, 3330     | 0.809 \[0.733-0.864\]     |      |        | 0.0025   | 65.4% |

<!-- numbers in the text updated up to here, CA 20/2/2025 -->

@@

Figure 3 presents (A) a forest and (B) funnel plot of the random-effects model of the best-performing models from all studies. Similarly to valence, the range of correlations is also wide for arousal, ranging from 0.35 to 0.95, with all studies demonstrating evidence of positive correlations. A mean estimate of 0.82 or higher is achieved by the majority (14 out of the 22 models). Due to large sample in most studies, the confidence intervals are narrow, although the exceptions ($N < 55$) are clearly visible [@beveridge2018po; @griffiths2021am;@koh2023me; @saizclar2022pr; @wang2021ac]. The funnel plot in panel B illustrates the limited range of the correlations and show heavy clustering at the top of the plot (studies with low standard error). There is no clear asymmetry in the plot, verified by non-significant Egger's test reported above.

The analysis of the subdivision of studies shows that there is no significant differences between the studies using different number of features (_Q_(2) = 5.20, _p_ = .074) despite the differing means (_r_ = 0.885 for studies with less than 30 features, _r_ = 0.735 for 30 to 300 features, and _r_ = 0.804 for studies utilising over 300 features). The differences in the techniques, however, does show statistically significant variance between subgroups (_Q_(3) = 10.83, _p_ = .0127). The Neural Nets (NN) achieve poor prediction of arousal (_r_ = 0.660) in comparison to other techniques. The caveat of this subgroup analysis is the small number of observations for four techniques. When other subgroupings were explored, such as the type of journal (Engineering with 13 studies vs Psychology with 9 studies), no significant differences were observed (_Q_(1) = 0.94, _p_ = .333).

{{< include _figure3.qmd >}}
<!-- numbers in the text updated up to here, CA 20/2/2025 -->
